\documentclass[10pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{amsmath,amssymb,booktabs,enumitem,hyperref,longtable,array}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\setlist{nosep,leftmargin=1.2em}
\setlength{\parskip}{0.35em}
\setlength{\parindent}{0pt}
\hypersetup{colorlinks=true,linkcolor=blue,urlcolor=blue}
\newcommand{\cls}[1]{\textbf{#1}}
\newcommand{\code}[1]{\texttt{#1}}

\title{\textbf{NBA Home-Win Probability Model}\\[0.25em]
\large Mathematics \& Full Constant Provenance\\[0.15em]
\normalsize Deployed champion: Elo-only (architecture \texttt{conservative}) --- technical briefing}
\author{Joseph Shackelford\\[0.15em]
\small Prepared for recruiter / technical review.\\
\small Every number below is classified
\cls{Fixed}, \cls{Selected}, \cls{Fitted}, \cls{Derived}, or \cls{Numerical},
with an explicit source or calculation.\\
\small Artifacts: \code{artifacts/selected\_spec.json},
\code{march\_architecture\_results.csv},
\code{coefficient\_table.csv}.}
\date{}

\begin{document}
\maketitle
\vspace{-1.0em}

\noindent\textit{Purpose of this document.}
This note states the model equations and accounts for \textbf{every numeric
constant}: if a number is calculated, the exact optimization is written; if it
is not calculated, the external or design source is named. No constant is left
as an unexplained ``magic number.''

\section{Champion model (one equation)}

The \textbf{deployed champion is Elo-only}: an L2-logistic map on the
standardized margin-of-victory Elo rating differential,
\[
\boxed{
p_g
=
\sigma\!\bigl(c + w\,\tilde x_g^{E}\bigr)
},
\qquad
\tilde x_g^{E}=\frac{x_g^{E}-\mu}{s},
\]
with \textbf{deployed} coefficients (fit on all games through 2026-03-31)
\(w=0.9271823510\),
\(c=0.2415428879\),
\(\mu=0.1408146877\),
\(s=0.2690252334\)
(raw-unit weight \(w/s\approx 3.4464512\)),
and
\(\sigma(z)=1/(1+e^{-z})\),
\(\operatorname{logit}(p)=\log(p/(1-p))\). This is exactly the calibrated Elo
component of \S\ref{sec:elo}; April performance state is frozen at 2026-03-31.

\paragraph{Rejected challenger.} An Elo \(+\) Bradley--Terry/recent-trend
convex logistic stacker,
\(p_g^{\mathrm{blend}}=\sigma(a\operatorname{logit}(p_g^{E})+b\operatorname{logit}(p_g^{R})+c)\)
with \(a,b\ge0,\ a+b\le1\), was implemented and \textbf{rejected}: under nested
rolling-origin validation (\S\ref{sec:nested}) it is worse than Elo-only on both
log loss and Brier and is worse calibrated. It is retained only in the
\code{challenger} block of \code{artifacts/selected\_spec.json}; see
\S\ref{sec:stacker}.

\section{How to read every constant}

\begin{center}
\begin{tabular}{@{}lp{4.6in}@{}}
\toprule
\textbf{Class} & \textbf{Meaning} \\
\midrule
\cls{Fixed} & Hardcoded design choice. Not estimated on this CSV. \\
\cls{Selected} & Chosen by comparing candidates on the \textbf{aggregate frozen-policy rolling out-of-sample surface} (one-standard-error stability rule). \\
\cls{Fitted} & Numeric solution of an optimization / MLE on data. \\
\cls{Derived} & Algebraic consequence of other constants (no new information). \\
\cls{Numerical} & Solver / clip / RNG setting; not a modeling claim. \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Honesty rule used below.}
If a number was not estimated or selected on this project’s data, this document
says so. In particular, the MOV constant \(2.2\) is \cls{Fixed} (literature): it
is \emph{not} re-fitted, but it \emph{is} profiled on the frozen-policy rolling
out-of-sample grid and kept because it sits within one standard error of the
grid best (\(3.0\)); see \S\ref{sec:mov}.

%==============================================================================
\section{Master ledger --- every constant}
%==============================================================================

\subsection{A. Elo core (\cls{Fixed} unless noted)}

\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.15in}
>{\raggedright\arraybackslash}p{0.72in}
>{\raggedright\arraybackslash}p{0.72in}
>{\raggedright\arraybackslash}p{3.55in}@{}}
\toprule
\textbf{Constant} & \textbf{Value} & \textbf{Class} & \textbf{Provenance / calculation} \\
\midrule
\endhead
Initial rating \(R_i(0)\) & \(1500\) & Fixed & Standard Elo origin. Code: \code{ratings = \{team: 1500.0\}} in \code{features.py}. Not fit. \\
Elo scale \(S\) & \(400\) & Fixed & Classical Elo logistic scale: a \(400\)-point gap \(\Rightarrow\) \(10{:}1\) odds in the raw Elo formula. Denominator in \(x^{E}=(R_h-R_a+H)/400\). \\
Base of Elo logistic & \(10\) & Fixed & Classical Elo uses base-\(10\) logistic, not \(e\). Formula \(1/(1+10^{-x})\). \\
MOV mode & \texttt{log} & Selected$^\star$ & Architecture field \code{elo\_mov}. All five candidates used \texttt{log}; mode itself was not raced against \texttt{none} in selection. \\
MOV blowout shape & \(\ln(|m|+1)\) & Fixed & FiveThirtyEight / Nate Silver NFL Elo log-MOV form. \\
MOV constant \(b_{\mathrm{MOV}}\) & \(2.2\) & Fixed (external) & Adopted from the published FiveThirtyEight NFL Elo MOV formula (Silver). \textbf{Not re-fitted on this CSV}, but profiled on the frozen-policy OOS grid \(\{1.6,2.0,2.2,2.6,3.0\}\) and confirmed within one SE of the grid best (\(3.0\)); recorded under \code{mov\_offset\_selection}. Full professional rationale, benefits, and the profile: \S\ref{sec:mov}. \\
MOV slope \(a_{\mathrm{MOV}}\) & \(0.001\) & Fixed & Paired with \(2.2\) in the same FTE formula. Ratio \(a/b\) sets favorite/underdog discount. Not fit here. \\
MOV denom floor & \(0.25\) & Fixed & Code guard \code{max(0.25, denom)} against pathological denominators. Not estimated. \\
\bottomrule
\end{longtable}
{\footnotesize $^\star$Selected only in the weak sense that the winning architecture kept \texttt{log}; no candidate used \texttt{none}.}

\subsection{B. Architecture hyperparameters for champion \texttt{conservative}}

These appear in \code{configs/architecture\_candidates.json} and the winner in
\code{artifacts/selected\_spec.json}.
\textbf{Selection rule:} the deployed Elo architecture is chosen by the
\textbf{aggregate frozen-policy rolling out-of-sample log loss} with a
\textbf{one-standard-error stability rule} (simplest / lowest-\(K\) architecture
within one SE of the best; \code{nba\_wp/selection.py},
\code{architecture\_selection} in \code{selected\_spec.json}). \texttt{conservative}
wins outright (lowest mean OOS log loss \emph{and} lowest \(K\)), and a single
March one-step split agrees. The MOV offset (\(2.2\)) and an early-season
provisional-\(K\) warmup (kept \emph{off}) are profiled on the same surface. The
champion is Elo-only; April rows loaded during selection: \(0\). The champion
depends only on the Elo fields \(K,H,\text{MOV},C_E\); the remaining fields govern
the rejected challenger.

\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.35in}
>{\raggedright\arraybackslash}p{0.55in}
>{\raggedright\arraybackslash}p{0.72in}
>{\raggedright\arraybackslash}p{3.5in}@{}}
\toprule
\textbf{Constant} & \textbf{Value} & \textbf{Class} & \textbf{Provenance / calculation} \\
\midrule
\endhead
\(K\) (Elo step) & \(7.5\) & Selected & Member of \texttt{conservative} candidate. Chosen by the one-SE stability rule on the aggregate frozen-policy OOS surface (lowest mean OOS log loss and lowest \(K\); Table~\ref{tab:march} shows the March cross-check). Not a continuous MLE. \\
\(H\) (home Elo pts) & \(55\) & Selected & Same bundle. Equal-team home edge \(\Rightarrow 1/(1+10^{-55/400})\approx 0.578\). \\
Elo logistic \(C_E\) & \(10\) & Selected & Regularization for the deployed Elo calibrator. From \texttt{conservative}. \\
\(C_{\mathrm{BT}}\) & \(0.1\) & Selected$^\dagger$ & Bradley--Terry regularization (challenger only). \\
Trend half-life \(\tau_{1/2}\) & \(60\) days & Selected$^\dagger$ & EWMA weight \(2^{-\Delta\mathrm{days}/60}\) (challenger only). \\
Trend short window \(n_S\) & \(12\) games & Selected$^\dagger$ & Mean of last \(\min(12,n_i)\) margins (challenger only). \\
Rank logistic \(C_R\) & \(0.1\) & Selected$^\dagger$ & BT+trend calibrator regularization (challenger only). \\
Architecture name & \texttt{conservative} & Selected & One-SE stability winner on the aggregate frozen-policy OOS surface over 5 named bundles (March split agrees). \\
\bottomrule
\end{longtable}
{\footnotesize $^\dagger$These fields do not enter the deployed Elo-only champion; they parameterize the rejected challenger blend.}

\subsection{C. Deployed Elo-only coefficients (\cls{Fitted})}
\label{sec:stacker}

The deployed champion has \textbf{no stacker}: it is the calibrated Elo logistic
(\S\ref{sec:elo}), fit on all rows through 2026-03-31. From
\code{artifacts/coefficient\_table.csv}:

\begin{longtable}{@{}llll@{}}
\toprule
\textbf{Component} & \textbf{Feature} & \textbf{Std.\ coef} & \textbf{Raw-unit coef} \\
\midrule
\endhead
Elo & \(x^{E}=\mathrm{elo\_diff}\) & \(0.92718235\) & \(3.44645125\) \\
Elo & intercept & \(0.24154289\) & --- \\
\bottomrule
\end{longtable}
Raw-unit coef \(=\) standardized coef \(/\) training scale
(\(s=0.26902523\), \(\mu=0.14081469\)). Fitted by penalized MLE (\code{lbfgs},
\(C_E=10\)).

\paragraph{Rejected challenger stacker.}
For completeness, the challenger blend's convex stacker (architecture
\texttt{conservative}) has \textbf{deploy} coefficients
\(a=0.3976558706\), \(b=0.6023441294\), \(c=0.3347604843\)
(\(a+b=1\), temperature \(\tau=1\)), from an unconstrained MLE fit
\(a=0.6076317564\), \(b=0.9204024091\), \(c=0.3279170537\)
(\(\tau=1/(a+b)\approx 0.655<1\)). \code{fit\_logit\_stacker} enforces a genuine
convex blend --- it clips any negative weight to zero and applies the
temperature floor \(\tau\ge1\), so \(0\le w\le1\) for both components (pinned by
\code{tests/test\_stacker\_temperature\_floor.py::\allowbreak test\_stacker\_weights\_are\_convex\_when\_stacker\_is\_used}).
This blend is not deployed; it loses to Elo-only out-of-sample
(\S\ref{sec:nested}). The full challenger coefficients live in the
\code{challenger} block of \code{artifacts/selected\_spec.json}.

\subsection{D. Challenger base-component coefficients (\cls{Fitted}, reference)}

The rejected challenger's rank component (Bradley--Terry \(+\) trend) is an
L2-logistic calibrator; its coefficients are archived with the challenger. They
are not part of the deployed champion, which uses only the Elo row in \S C.

\noindent\textbf{How calculated.} For training rows available at that stage,
\[
\hat{\boldsymbol{\beta}}
=
\arg\max_{\boldsymbol{\beta}}
\sum_i
\big[
Y_i\log\sigma(\beta_0+\boldsymbol{\beta}^\top\tilde{\mathbf{x}}_i)
+(1-Y_i)\log(1-\cdots)
\big]
-\tfrac{1}{2C}\|\boldsymbol{\beta}\|_2^2,
\]
with \(\tilde{\mathbf{x}}\) standardized (\code{StandardScaler}),
\(C=C_E\) or \(C_R\), solver \code{lbfgs}.
Raw-unit coef \(=\) standardized coef \(/\) training scale.

\subsection{E. Bradley--Terry / trend auxiliaries}

\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.35in}
>{\raggedright\arraybackslash}p{0.7in}
>{\raggedright\arraybackslash}p{0.72in}
>{\raggedright\arraybackslash}p{3.35in}@{}}
\toprule
\textbf{Constant} & \textbf{Value} & \textbf{Class} & \textbf{Provenance} \\
\midrule
\endhead
BT design coding & \(+1/-1\) & Fixed & Home column \(+1\), away \(-1\). Standard BT contrast. \\
BT team strengths \(q_i,\alpha\) & (vector) & Fitted & L2-logistic MLE on all prior games before date \(d\), \(C=C_{\mathrm{BT}}\). Refit when history grows. Feature \(x^{\mathrm{BT}}=\hat\alpha+\hat q_h-\hat q_a\). \\
EWMA base \(1/2\) & \(0.5\) & Fixed & Half-life definition: weight \(2^{-\Delta/\tau_{1/2}}=(1/2)^{\Delta/\tau_{1/2}}\). \\
Beta prior \((\alpha,\beta)\) & \((4,4)\) & Fixed & Comment in code: ``eight games of neutral information.'' Smoothed win rate \((W+4)/(G+8)\). Used for \code{record\_logit} (ablation feature; not in champion \(X\)). \\
Rest cap & \(7\) days & Fixed & Rest days clipped to \([0,7]\). Schedule features only (ablation path). \\
B2B threshold & \(1\) day & Fixed & \code{back\_to\_back = (rest <= 1)}. Ablation path. \\
Density windows & \(4\), \(6\) days & Fixed & Games-in-window counts. Ablation path. \\
\bottomrule
\end{longtable}

\subsection{F. Calendar / selection policy (\cls{Fixed} policy)}

\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.5in}
>{\raggedright\arraybackslash}p{1.35in}
>{\raggedright\arraybackslash}p{0.72in}
>{\raggedright\arraybackslash}p{2.55in}@{}}
\toprule
\textbf{Constant} & \textbf{Value} & \textbf{Class} & \textbf{Provenance} \\
\midrule
\endhead
Train-through (base, March scoring) & 2026-02-28 & Derived & Derived from the data (\code{nba\_wp/periods.py}), not hard-coded. \\
Selection window & 2026-03-01--03-31 & Derived & Month before the holdout (\code{derive\_periods}). \\
April holdout start & 2026-04-01 & Derived & Last calendar month in the data (\code{derive\_periods}); selection raises if April present. \\
Accuracy cutoff & \(0.5\) & Fixed & \code{p >= 0.5} in metrics. Not optimized. \\
Primary metric & log loss & Fixed policy & Selection rule text. \\
\# architectures raced & \(5\) & Fixed design & Length of candidate list. \\
\bottomrule
\end{longtable}

\subsection{G. Derived quantities (\cls{Derived})}

\begin{longtable}{@{}>{\raggedright\arraybackslash}p{1.2in}
>{\raggedright\arraybackslash}p{0.85in}
>{\raggedright\arraybackslash}p{4.1in}@{}}
\toprule
\textbf{Quantity} & \textbf{Value} & \textbf{Calculation} \\
\midrule
\endhead
Elo raw-unit weight & \(3.4465\) & \(w/s=0.92718235/0.26902523\) (deployed champion) \\
Challenger \(w=a/(a+b)\) (raw) & \(0.3977\) & \(0.60763176/(0.60763176+0.92040241)\); preserved by the floor \\
Challenger \(\tau\) (raw) & \(0.6551\) & \(1/(0.60763176+0.92040241)<1\Rightarrow\) floor applied \\
Challenger \(\tau\) (deploy) & \(1.0000\) & \(1/(0.39765587+0.60234413)\) after projection onto \(a+b=1\) \\
Equal-team home \(p\) & \(\approx 0.578\) & \(1/(1+10^{-H/400})\) with \(H=55\) \\
Fair home odds & \(1/p_g\) & Definition \\
Fair away odds & \(1/(1-p_g)\) & Definition \\
\bottomrule
\end{longtable}

\subsection{H. Numerical / solver settings (\cls{Numerical})}

\begin{longtable}{@{}lll@{}}
\toprule
\textbf{Setting} & \textbf{Value} & \textbf{Where} \\
\midrule
\endhead
\code{max\_iter} & \(10000\) & all logistics \\
\code{tol} & \(10^{-12}\) & all logistics \\
\code{random\_state} & \(0\) & reproducibility \\
\code{solver} & \code{lbfgs} & all logistics \\
probability clip (logit) & \([10^{-12},1-10^{-12}]\) or \([10^{-9},1-10^{-9}]\) & \code{model.py}/\code{features.py} \\
\bottomrule
\end{longtable}

%==============================================================================
\section{Target and information}
%==============================================================================

\[
Y_g=
\begin{cases}
1,&\text{home points}_g>\text{away points}_g,\\
0,&\text{otherwise.}
\end{cases}
\qquad
X_g=f(\mathcal{H}_{<d}).
\]
Same-date batching: write all rows for date \(d\) before any Elo/BT/trend update
from results on \(d\).

%==============================================================================
\section{Elo mathematics (the deployed champion)}
\label{sec:elo}
%==============================================================================

\subsection{Raw probability and feature}
\[
p_g^{\mathrm{Elo,raw}}
=
\frac{1}{1+10^{-(R_h-R_a+H)/400}},
\qquad
x_g^{E}
=
\frac{R_h-R_a+H}{400}.
\]
Constants: \(R_i(0)=1500\) (Fixed), \(400\) (Fixed), \(H=55\) (Selected).

\subsection{Update with MOV}
\[
R_h' = R_h + K M_g (Y_g-p_g^{\mathrm{Elo,raw}}),\qquad
R_a' = R_a - K M_g (Y_g-p_g^{\mathrm{Elo,raw}}),
\]
\[
M_g
=
\ln(|m_g|+1)
\cdot
\frac{2.2}{\max\bigl(0.25,\,0.001(R_{\mathrm{win}}-R_{\mathrm{lose}})+2.2\bigr)},
\]
where \(R_{\mathrm{win}},R_{\mathrm{lose}}\) are the pregame ratings of the
\emph{winning} and \emph{losing} team (\S\ref{sec:movsign}). Constants:
\(K=7.5\) (Selected); \(2.2\), \(0.001\), \(0.25\) (Fixed).

\subsection{MOV rating-difference sign (winner \(-\) loser)}
\label{sec:movsign}
The rating difference inside the MOV denominator is the \emph{winner minus
loser} rating, not a fixed \emph{home minus away} rating. This matches the
published FTE form \((\mathrm{ELO}_W-\mathrm{ELO}_L)\) and has two correct
consequences: (i) an \textbf{upset} (a lower-rated team winning) gives a
smaller denominator and hence a \emph{larger} rating update than an expected
result, and (ii) with zero home advantage the update is exactly team-swap
symmetric. Using a fixed home\(-\)away difference instead would make the
autocorrelation correction depend on venue rather than on who actually won,
under-updating home upsets and over-updating home favorites. This behaviour is
pinned by \code{tests/test\_elo\_mov\_winner\_diff.py}.

\subsection{The constant \(2.2\): source, case for use, and how to optimize it}
\label{sec:mov}

\paragraph{Disclosure (external source).}
The value \(2.2\) was \textbf{not estimated from this project's NBA dataset}.
It was adopted from a published, widely cited sports-rating specification:
Nate Silver / FiveThirtyEight,
\textit{Introducing NFL Elo Ratings}
(archived:
\url{https://web.archive.org/web/20171004105151/https://fivethirtyeight.com/features/introducing-nfl-elo-ratings/}),
which defines
\[
M
=
\ln(|PD|+1)
\cdot
\frac{2.2}{(\mathrm{ELO}_W-\mathrm{ELO}_L)\cdot 0.001 + 2.2}.
\]
This submission copies that functional form into
\code{nba\_wp/features.py::\_elo\_multiplier}
when \code{elo\_mov="log"}. Transparency on this point is intentional:
industry models routinely reuse validated structural constants when the
\emph{form} is what matters and the local sample is too short to re-estimate
every shape parameter cleanly.

\paragraph{Professional case for using \(2.2\) here (best case, optimal or not).}
Even though \(2.2\) is external, using this specific published constant is a
defensible engineering choice for a leakage-audited pricing model:

\begin{enumerate}
\item \textbf{Correct economic shape, not an arbitrary fudge.}
The log term gives diminishing returns to blowouts (a 30-point win should not
update ratings like three independent 10-point wins). The fraction involving
\(2.2\) implements the standard underdog/favorite correction so favorites that
``run up the score'' do not inflate permanently. That is precisely the
behavior a book wants in a strength tracker feeding probabilities.
\item \textbf{Stability and reproducibility.}
A public constant with a fixed formula is auditable: two reviewers regenerating
features get the same Elo path. Replacing it with an in-sample tuned value on
one partial season would add a free parameter, widen the selection surface, and
invite overfit to March noise.
\item \textbf{Separation of concerns in the validation design.}
This project concentrates selection effort on quantities that move prices most
(home advantage \(H\), step size \(K\), BT/trend regularization) under a rolling
out-of-sample stability rule, while holding the MOV \emph{shape} at a known
industry default and profiling only its offset (which the one-SE rule keeps at
\(2.2\)). That is a standard bias--variance trade: fix low-sensitivity shape
parameters; select high-sensitivity ones.
\item \textbf{Empirical support that the choice is not fragile.}
Varying the constant in a neighborhood of \(2.2\) changes the Elo-only March log
loss only in the fourth decimal place (flat likelihood). So the \emph{presence}
of this MOV correction is beneficial; the exact published \(2.2\) is a stable
operating point, not a knife-edge claim of global optimality.
\item \textbf{Interpretability for stakeholders.}
\(2.2\) has a clear reading in the literature as the baseline expected
\(\ln(|PD|+1)\) scale in the autocorrelation correction (see Morse's
derivation of \(d=2200=2.2/0.001\)). That is preferable to an opaque
tuned scalar with no external meaning.
\end{enumerate}

\paragraph{How the offset is profiled (now implemented).}
This is no longer a thought experiment: \code{nba\_wp/selection.py::profile\_mov\_offset}
profiles the offset on the same \textbf{frozen-policy rolling out-of-sample surface}
used for architecture selection, and keeps the literature value \(2.2\) unless a
grid value beats it beyond one standard error (recorded under
\code{mov\_offset\_selection} in \code{selected\_spec.json}). If instead one wanted
the raw \emph{data-optimal} constant \(b^\star\), compute it by the same principle
while holding the champion architecture fixed:

\begin{enumerate}
\item Fix architecture \texttt{conservative} except replace the literal \(2.2\) by a
candidate \(b>0\) in
\[
M_g(b)
=
\ln(|m_g|+1)
\cdot
\frac{b}{\max\bigl(0.25,\,0.001(R_h-R_a)+b\bigr)}.
\]
(Optionally keep the literature ratio by setting the slope to \(b/2200\) instead
of \(0.001\); the one-parameter profile below is the minimal change.)
\item For each candidate \(b\) on a grid
(e.g.\ \(b\in\{1.0,1.5,\ldots,4.0\}\)) or by continuous search:
build features \(\rightarrow\) fit base models on games before 2026-03-01
\(\rightarrow\) fit the March stacker \(\rightarrow\) record March log loss
\(\mathrm{LL}(b)\).
\item Set the optimal replacement
\begin{center}
\fbox{\parbox{0.88\linewidth}{\centering\vspace{0.55em}
\(
\displaystyle
b^{\star}
\;=\;
\operatorname*{arg\,min}_{b}\;
\mathrm{LL}_{\mathrm{March}}(b)
\)
\vspace{0.55em}}}
\end{center}
On the committed frozen-policy OOS grid \(b\in\{1.6,2.0,2.2,2.6,3.0\}\) the
\emph{nominal} minimizer is \(b^{\star}=3.0\) (mean OOS LL \(0.54174\)), but \(2.2\)
(mean OOS LL \(0.54204\)) sits well within one standard error of it
(SE \(\approx 0.019 \gg \Delta\mathrm{LL}\approx 3\times10^{-4}\)). Hence the
one-standard-error rule \textbf{keeps \(2.2\)}: it is not claimed as \(b^{\star}\),
but as a published, stable default that is statistically tied with the grid
optimum on our data.
\item \textbf{Stronger production estimator (multi-season).}
On a longer NBA history, estimate the literature intercept directly:
regress bin-wise
\(\widehat{\mathbb{E}}[\ln(|PD|+1)\mid\mathrm{win},\,x]\)
on Elo gap \(x\) and take the intercept \(\hat b_0\) (and slope \(\hat a_0\))
as in the Morse/FTE construction, then set \(b\leftarrow\hat b_0\).
Alternatively, nest \(b\) inside the architecture search and select by
prequential log loss with a locked final holdout.
\end{enumerate}

\paragraph{One-sentence position for reviewers.}
\textit{I used the published FiveThirtyEight log-MOV constant \(2.2\) as a fixed,
auditable shape parameter because it encodes a validated blowout/underdog
correction, keeps the Elo path reproducible, and sits on a flat region of the
March likelihood; a March (or multi-season) profile of \(\mathrm{LL}(b)\) is the
correct way to replace it with a data-optimal \(b^\star\) if that becomes a
production requirement.}

\subsection{Calibrated Elo probability (the deployed price)}
\[
p_g^{E}
=
\sigma(\beta_0^{E}+\beta_1^{E}\tilde x_g^{E})
= p_g,
\quad C_E=10\ \text{(Selected)},
\]
with deployed \(\beta_0^{E}=0.2415428879\), \(\beta_1^{E}=0.9271823510\)
(standardized). This \emph{is} the champion of \S1.

%==============================================================================
\section{Bradley--Terry + trend (rejected challenger only)}
%==============================================================================

The following components feed only the rejected challenger blend, not the
deployed Elo-only champion.

\subsection{Bradley--Terry}
Prior games before date \(d\):
\[
P(Y=1)=\sigma(\alpha+q_h-q_a),
\quad C_{\mathrm{BT}}=0.1\ \text{(Selected)}.
\]
Feature: \(x_g^{\mathrm{BT}}=\hat\alpha+\hat q_h-\hat q_a\) (\cls{Fitted} each day).

\subsection{Trend}
\[
L_{i,d}
=
\frac{\sum_{k<d} 2^{-\Delta_{kd}/60} m_{ik}}{\sum_{k<d} 2^{-\Delta_{kd}/60}},
\qquad
S_{i,d}
=
\frac{1}{n_i^{\mathrm{short}}}\sum_{\mathrm{last}\,n_i^{\mathrm{short}}} m_{ik},
\]
\[
n_i^{\mathrm{short}}=\min(12,n_i),
\quad
T_{i,d}=S_{i,d}-L_{i,d},
\quad
x_g^{\mathrm{trend}}=T_{h,d}-T_{a,d}.
\]
Constants \(60\) and \(12\): \cls{Selected} with \texttt{conservative}.
Base \(2\) in \(2^{-\cdot}\): \cls{Fixed} half-life algebra.

\subsection{Rank component}
\[
p_g^{R}
=
\sigma(
\beta_0^{R}
+\beta_{\mathrm{BT}}\tilde x_g^{\mathrm{BT}}
+\beta_{\mathrm{tr}}\tilde x_g^{\mathrm{trend}}
),
\quad C_R=0.1\ \text{(Selected)}.
\]

%==============================================================================
\section{Selection proof for architecture constants}
\label{tab:march}

\textbf{Elo-only} March sequential log loss (base fit through February) from
\code{artifacts/march\_architecture\_results.csv} (lower LL wins). This single
March split is a \emph{cross-check}: the deployed architecture is actually chosen
by the aggregate frozen-policy rolling OOS one-standard-error stability rule
(\code{architecture\_selection} in \code{selected\_spec.json}), and it agrees with
this table:

\begin{center}
\begin{tabular}{@{}lcccc@{}}
\toprule
Architecture & \(K\) & \(H\) & Elo-only March LL & Winner \\
\midrule
\textbf{conservative} & \(7.5\) & \(55\) & \(\mathbf{0.50659003}\) & yes \\
hfa\_75 & \(10\) & \(75\) & \(0.50909006\) & \\
balanced & \(10\) & \(65\) & \(0.50913895\) & \\
lower\_bt\_shrinkage & \(10\) & \(65\) & \(0.50913895\) & \\
responsive & \(15\) & \(65\) & \(0.51470876\) & \\
\bottomrule
\end{tabular}
\end{center}

\noindent \(K=7.5\), \(H=55\), \(C_E=10\) are the coordinate values of the
\texttt{conservative} bundle, which is the one-SE stability winner on the
aggregate frozen-policy OOS surface (and the minimum of this March cross-check).
The MOV offset (\(2.2\)) and provisional-\(K\) warmup (off) \emph{are} profiled on
the same OOS surface (\code{mov\_offset\_selection}, \code{cold\_start\_selection}
in \code{selected\_spec.json}). (Each other procedure additionally selects its own
architecture; see \code{march\_architecture\_results.csv}.)

%==============================================================================
\section{Metrics (definitions; cutoffs Fixed)}
%==============================================================================

\[
\mathrm{LL}
=
-\frac1N\sum_g
\big[Y_g\log p_g+(1-Y_g)\log(1-p_g)\big],
\quad
\mathrm{BS}
=
\frac1N\sum_g(p_g-Y_g)^2.
\]
Accuracy uses Fixed cutoff \(p_g\ge 0.5\).

Champion results (Elo-only, artifacts). \textbf{Primary April holdout (frozen
pre-April)}: LL \(0.464369\), Brier \(0.149770\), AUC \(0.866847\),
accuracy \(78.13\%\) (\(N=96\)). March selection period (Elo-only, base fit
through February, \(N=239\)): LL \(0.506590\). Optional April sequential
backtest (live-update simulation): LL \(0.464648\). Rejected blend on the same
frozen April window: LL \(0.468725\), Brier \(0.150465\) --- worse on both.

%==============================================================================
\section{Nested rolling-origin validation (why Elo-only)}
\label{sec:nested}
%==============================================================================

\code{scripts/nested\_validation.py} validates the whole procedure under two
information policies (frozen-block and daily-sequential; never mixed), 11 weekly
outer folds, 501 out-of-sample games. Each procedure independently selects its
own architecture by its own inner out-of-fold score; the challenger stacker
trains on inner out-of-fold predictions. Pooled out-of-sample (frozen-block):
Elo-only LL \(0.532\)/Brier \(0.177\); rank-only \(0.550\)/\(0.184\); blend
\(0.548\)/\(0.183\); constant \(0.688\)/\(0.247\). Block-bootstrap
blend\(-\)Elo: \(\Delta\mathrm{LL}=+0.017\) (95\% CI \([+0.010,+0.023]\)),
\(\Delta\mathrm{Brier}=+0.006\) (95\% CI \([+0.004,+0.009]\)); \textbf{0 of
4{,}000} week-block replicates favored the blend. Two further challengers are
added to the same audit and \textbf{also rejected} under both policies: a
cross-fitted \textbf{calibrated Elo} (identity-shrunk \(\alpha+\beta\operatorname{logit}(p_E)\);
over-corrects out-of-sample, \(\beta\approx1.79\); \(0.548/0.183\)) and a
\textbf{schedule Elo} (Elo + rest + back-to-back, regularized; \(0.531/0.177\),
slightly worse on both scores). Calibration (frozen-block): Elo-only slope
\(\beta\approx1.37\) (95\% CI \([1.22,1.57]\)), ECE \(\approx 0.059\) --- better
calibrated than the blend (\(\beta\approx1.80\)). Decision under both policies:
\textbf{keep Elo-only}. Artifacts:
\code{nested\_frozen\_block\_summary.json},
\code{nested\_daily\_sequential\_summary.json}.

%==============================================================================
\section{End-to-end constant flow}
%==============================================================================

Deployed champion (Elo-only):
\[
\underbrace{1500,400,2.2,0.001}_{\text{Fixed Elo}}
\;+\;
\underbrace{K,H,C_E}_{\text{Selected}}
\;\xrightarrow{\text{features}}\;
x^{E}
\;\xrightarrow[\text{Fitted }]{c,\,w}\;
p^{E}=p.
\]

\noindent\textbf{Code map.}
\code{nba\_wp/features.py} (Fixed Elo/MOV; Selected architecture fields),
\code{nba\_wp/model.py} (Fitted Elo \(c,w\); challenger \(a,b,c\)),
\code{nba\_wp/selection.py} (aggregate frozen-policy OOS one-SE architecture selection; MOV/cold-start profiling),
\code{artifacts/selected\_spec.json} (persisted Elo-only champion + rejected challenger).

\end{document}
